---
title: "dplyr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

#reading the dataset(please enter the path where you save the data)
wine <- read.csv("E:/Extra/Meet Up/winemag-data_first150k.csv",stringsAsFactors = F,encoding = "UTF-8")
```

# installing and loading dplyr package

```{r message=FALSE, warning=FALSE}
# install.packages("dplyr")
library(dplyr)
library(knitr)
```

# dplyr package

- dplyr is a powerful R-package to transform and summarize tabular data 
- 6 dplyr functions are,
  - select() - select columns
  - filter() - filter rows
  - arrange() - re-order or arrange rows
  - summarise() - summarise values
  - group_by() - allows for group operations in the "split-apply-combine"    concept
  - mutate() - create new columns


# 1. select() function

- used to select certain particular columns from the dataset

#selecting few columns with respect to index and top 10 rows 

```{r}
kable(wine %>% 
        select(1,2) %>% 
        head(10))
```

#removing 1st and 3rd column from the dataset

```{r}

wine = wine %>% 
  select(-c(1,3)) %>% 
  head()
```

#selecting columns with respect to their names and botton 6 rows

```{r}
kable(wine %>% 
        select(country) %>% 
        tail())
```


# 2. filter() function

- used to filter data with reapect to certain value


#filtering data for only conunty US 

```{r}
kable(wine %>% 
        filter(country == "US") %>% 
        head(10))
```

#filtering data for price greater than 100$

```{r}
kable(wine %>% 
        filter(price > 100) %>% 
        head(10))
```

#creating a vector of selected country
#filtering with respect to the values in the vector

```{r}
country_selected = c("US","France")

kable(wine %>% 
        filter(country %in% country_selected) %>% 
        head(10))

```


# 3. arrange() function

- used to arrange the data with respect to the values in particular column
- data can be arranged either in Ascending or Descending order


#arranges the selected columns in ascending order of points

```{r}
kable(wine %>% 
        arrange(points) %>% 
        select(country,points) %>% 
        head(10))
```

#arrange the selected columns in descending order of price

```{r}
kable(wine %>% 
        arrange(desc(price)) %>% 
        select(country,price) %>% 
        head(10))
```

#one more way to arrange the data in descending order

```{r}
kable(wine %>% 
        arrange(-price) %>% 
        select(country,price) %>% 
        head(10))

```


# 4. summarise() function

- output of summarise() is a value
- we can use any aggregate function inside summarise() to perform operations


#finding the average wine price

```{r}
kable(wine %>% 
        summarise(avg_price = mean(price,na.rm = T)))
```

#finding the maximum price of wine

```{r}
kable(wine %>% 
        summarise(max_price = max(price,na.rm = T)))
```

#finding the minimum price of wine

```{r}
kable(wine %>% 
        summarise(max_price = min(price,na.rm = T)))

```


# 5. group_by() function

- used to group data with respect to a value in the factor column
- this can be helpful to perform aggregate operations on a group
- its nothing but combinding similar data together to perform some operation

#grouping the countries together and finding their mean wine prices

```{r}
kable(wine %>% 
        group_by(country) %>% 
        summarise(avg_price = mean(price,na.rm = T)) %>% 
        head(10))
```

#grouping by variety and finding the average prices

```{r}
kable(wine %>% 
        group_by(variety) %>% 
        summarise(avg_price = mean(price,na.rm = T)) %>% 
        head(10))


```


# 6. mutate() function

- used to create a new column in the dataset

# creating a new coulmn called the ratio = points/price

```{r}
kable(wine %>% 
        mutate(ratio = points/price) %>% 
        head(10))

```



### total number of wines per country
### grouping by countries and then finding the count 

```{r message=FALSE, warning=FALSE}
countries_count = wine %>% 
  group_by(country) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```


###top 10 mass producing countries
###grouping by countries and then finding the count ,arranging them in decending order and finding the top 10 countries

```{r}
selected_countries = wine %>% 
  group_by(country) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  top_n(10) %>% 
  select(country)

kable(selected_countries)

kable(class(selected_countries))
```


###converting the single column data frame into a vector

```{r}
selected_countries = as.character(selected_countries$country)
kable(class(selected_countries))
```


###average points of top 10 mass producing countries
###filtering top 10 countries and displaying their average points

```{r}
select_avg_point = wine %>%
  group_by(country) %>% 
  filter(country %in% selected_countries) %>% 
  summarize(avg_points = mean(points)) %>%
  arrange(desc(avg_points)) %>% head(10)
kable(select_avg_point)

```

# points of top 10 mass producing countries
# filtering top 10 countries and displaying their points

```{r}
select_point = wine %>% 
  filter(country %in% selected_countries) %>% 
  select(country,points) %>% 
  arrange(country)
kable(select_point)
```


# Other top 10 producers of large quantity but low quality

```{r}

kable(wine %>% 
        filter(!(country %in% selected_countries)) %>% 
        group_by(country) %>% 
        summarise(median = median(points))%>% 
        arrange(desc(median)))
```


# Top countries with high points 

```{r}
top = wine %>% 
  group_by(country) %>% 
  summarise(median = median(points)) %>% 
  arrange(desc(median))

kable(top)
top = as.character(top$country)
```


# Grouping wrt the type of grapes

```{r}
topwine = wine %>% 
  group_by(variety) %>% 
  summarise(number = n()) %>% 
  arrange(desc(number)) %>% 
  head(10)

kable(topwine)
```


